---
title: "Tabula muris clustering and pseduotime"
description: |
  I did the Aaron Lun pipeline for this 10x data but now I want to go a bit more off the beaten path
author:
  - name: Daniel Brown
date: "`r Sys.Date()`"
output: 
  radix::radix_article:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup}
require("Matrix")
library(here)
library(DropletUtils)
library(scater)
library(scran)
library(SC3)
library(monacle)
library(EnsDb.Mmusculus.v79)
#setwd("/Volumes/SlowDrive/NGSdata/tenX_genomics/tabulaMuris/")
sce <- readRDS(here("data", "tabularMuris_withMarkers.rds"))
sce
knitr::opts_chunk$set(fig.path = "../plots/")
```
I notice there are cell labels stored in the orginal dataset lets look clusterings on cell type
```{r}
par(mfrow=c(1,2))
plotPCA(sce, ncomponents=2, colour_by="type")
plotUMAP(sce, colour_by="type")
plotTSNE(sce, colour_by="type")
```


Let’s run SC3 clustering on the Tabula Muris data. The advantage of the SC3 is that it can directly ingest a SingleCellExperiment object.
# SC3
## Estimation
```{r}
sce <- sc3_estimate_k(sce)
metadata(sce)$sc3$k_estimation
```
SC3 estimates 24 which I think is close to what SNN graph based clustering ID'd.
## Now the full SC3 classification
```{r}
# Got to write the gene names to a different slot
rowData(sce)$feature_symbol = row.names(sce)
sce <- sc3(sce, ks = 24, biology = TRUE, rand_seed = 10)
```
The run time therre was very long. Look at the summary plots
```{r}
sc3_plot_consensus(sce, k = 24, show_pdata = "type")
```
# Psudotime ordering

Monocle directly builds a minimum spanning tree on a reduced dimension representation of the cells to connect all cells. Monocle then identifies the longest path in this tree to determine pseudotime.
Unfortunately, Monocle does not work when all the genes are used, so we must carry out feature selection. First, we genes without counts in any cell.
```{r}
keep_feature <- rowSums(counts(sce) > 0) > 0
sce <- sce[keep_feature, ]
```
FIrst we need to define the genes used in the ordering.
I will try use all genes as a toy example
```{r}
rowData(sce)$gene_short_name = row.names(sce)
mon = convertTo(sce, type="monocle",row.fields=c(1,4,5,6,11), 
                col.fields=c(1,2,5))
mon <- estimateDispersions(mon)
HSMM_marrow <- setOrderingFilter(mon, row.names(sce))
plot_ordering_genes(HSMM_marrow)
```
I think the zeros in the data caused a problem for the dispersion estimate
```{r}
mon <- reduceDimension(mon, pseudo_expr = 1)
mon <- orderCells(mon, reverse = FALSE)
plot_cell_trajectory(mon)
```

