---
title: "Tabula muris clustering and pseduotime"
description: |
  I did the Aaron Lun pipeline for this 10x data but now I want to go a bit more off the beaten path
author:
  - name: Daniel Brown
date: "`r Sys.Date()`"
output: 
  radix::radix_article:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup}
require("Matrix")
library(here)
library(DropletUtils)
library(scater)
library(scran)
library(SC3)
library(EnsDb.Mmusculus.v79)
#setwd("/Volumes/SlowDrive/NGSdata/tenX_genomics/tabulaMuris/")
sce <- readRDS(here("data", "tabula", "tabularMuris_withMarkers.rds"))
sce
knitr::opts_chunk$set(fig.path = "../plots/")
```
I notice there are cell labels stored in the orginal dataset lets look clusterings on cell type
```{r}
par(mfrow=c(1,2))
plotPCA(sce, ncomponents=2, colour_by="type")
plotUMAP(sce, colour_by="type")
plotTSNE(sce, colour_by="type")
```


Let’s run SC3 clustering on the Tabula Muris data. The advantage of the SC3 is that it can directly ingest a SingleCellExperiment object.
# SC3
## Estimation
```{r}
sce <- sc3_estimate_k(sce)
metadata(sce)$sc3$k_estimation
```
SC3 estimates 24 which I think is close to what SNN graph based clustering ID'd.
## Now the full SC3 classification
```{r}
# Got to write the gene names to a different slot
rowData(sce)$feature_symbol = row.names(sce)
sce <- sc3(sce, ks = 24, biology = TRUE, rand_seed = 10)
```
The rrun time therre was very long. Look at the summary plots
```{r}
sc3_plot_consensus(sce, k = 24, show_pdata = "type")
```

