---
title: "Multi-sample comparisons"
author: "D_Brown"
date: "25/02/2020"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data import
```{r}
library(MouseGastrulationData)
sce.chimera <- WTChimeraData(samples=5:10)
```

# Feature-annotation

```{r}
library(scater)
rownames(sce.chimera) <- uniquifyFeatureNames(
    rowData(sce.chimera)$ENSEMBL, rowData(sce.chimera)$SYMBOL)
```

# quality-control
already performed by authors

```{r}
# We also remove cells labelled as stripped nuclei or doublets
drop <- sce.chimera$celltype.mapped %in% c("stripped", "Doublet")
sce.chimera <- sce.chimera[,!drop]
```

# normalization
using pre-computed size factors.

```{r}
sce.chimera <- logNormCounts(sce.chimera)
```

# variance-modelling
Retaining all genes with a positive biological component, to  preserve as much signal as possible in a very heterogeneous dataset.

```{r}
library(scran)
dec.chimera <- modelGeneVar(sce.chimera, block=sce.chimera$sample)
chosen.hvgs <- dec.chimera$bio > 0
```

# merging
hierarchical merge, first merge together reps with the # same genotype, then merge samples across different genotypes.

```{r}
library(batchelor)
set.seed(01001001)
merged <- correctExperiments(sce.chimera,
  batch=sce.chimera$sample,
  subset.row=chosen.hvgs,
  PARAM=FastMnnParam(
    merge.order=list(
      list(1,3,5), # WT (3 replicates)
      list(2,4,6)  # td-Tomato (3 replicates)
    )
  )
)
```

# clustering

```{r}
g <- buildSNNGraph(merged, use.dimred="corrected")
clusters <- igraph::cluster_louvain(g)
merged$cluster <- factor(clusters$membership)
```

# dimensionality-reduction
Using an external algorithm to compute nearest  neighbors for greater speed.  

```{r}
merged <- runTSNE(merged, dimred="corrected",
  external_neighbors=TRUE)
merged <- runUMAP(merged, dimred="corrected",
  external_neighbors=TRUE)

plotReducedDim(merged, colour_by = "celltype.mapped", dimred = "UMAP")
```

# Differential abundance
```{r}
abundances <- table(merged$celltype.mapped, merged$sample)
abundances <- unclass(abundances)
```

## Attaching some column metadata
```{r}
extra.info <- colData(merged)[match(
  colnames(abundances), merged$sample),]
y.ab <- DGEList(abundances, samples=extra.info)
```

## Filter out low-abundance labels
```{r}
keep <- filterByExpr(y.ab, group=y.ab$samples$tomato)
y.ab <- y.ab[keep,]
```

# Construct design matrix,
Blocking on pool and  including term for td-Tomato-status

```{r}
design <- model.matrix(
  ~0 + factor(pool) + factor(tomato),
  y.ab$samples)
# Tidy up design matrix column names
colnames(design) <- gsub(
  "factor|\\(|\\)|TRUE", "", colnames(design))
```

## Estimate and plot the negative binomial dispersions

```{r}
y.ab <- estimateDisp(y.ab, design, trend="none")
plotBCV(y.ab, cex=1)

# Estimate and plot the quasi-likelihood dispersions
fit.ab <- glmQLFit(y.ab, design, robust=TRUE,
  abundance.trend=FALSE)
plotQLDisp(fit.ab, cex=1)
```

# Test for DA due to td-Tomato-status
```{r}
res <- glmQLFTest(fit.ab, coef="tomato")
# Summarise the results
summary(decideTests(res))

# View the top DE genes
datatable(as.data.frame(
  topTags(res)
  ))
```
